### 2022-09-12 analysis.Rmd
update

---
title: "Face Attractiveness and the Sense of Reality"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
options(
  digits = 3,
  mc.cores = 4,
  brms.algorithm = "sampling",
  brms.backend = "cmdstanr"
)

cache <- TRUE
runModels <- FALSE
bestModels <- FALSE
fig.width <- see::golden_ratio(7)
fig.height <- 7

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.path = "./figures/",
  fig.width = fig.width,
  fig.height = fig.height
)
```

# Methods

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(brms)

df <- read.csv("data/data.csv") |>
  mutate(
    Date = lubridate::dmy(Date),
    Participant = fct_reorder(Participant, Date),
    Screen_Refresh = as.character(Screen_Refresh),
    Education = fct_relevel(Education, "Doctorate", "Master", "Bachelor", "High School", "Other", "Prefer not to Say"),
    Belief = fct_relevel(Belief, "Fake", "Real"),
    Stimulus_Interest = case_when(
      Sexual_Orientation == "Heterosexual" & Stimulus_SameSex == "Opposite" ~ TRUE,
      Sexual_Orientation == "Heterosexual" & Stimulus_SameSex == "Same" ~ FALSE,
      Sexual_Orientation == "Homosexual" & Stimulus_SameSex == "Opposite" ~ FALSE,
      Sexual_Orientation == "Homosexual" & Stimulus_SameSex == "Same" ~ TRUE,
      Sexual_Orientation %in% c("Bisexual", "Queer") ~ TRUE,
      TRUE ~ NA
    )
  )

# head(df[is.na(df$Stimulus_Attract), ])

# Create individual scores for Simulation Monitoring
df <- df |>
  group_by(Participant, Belief) |>
  summarise(
    Confidence = mean(abs(Belief_Confidence)),
    n = n() / 109
  ) |>
  pivot_wider(names_from = "Belief", values_from = c("Confidence", "n")) |>
  ungroup() |>
  merge(df, by = "Participant")
```

## Exclusions {.tabset}

```{r message=FALSE, warning=FALSE}
outliers <- c(
  # Very short duration for questionnaire in particular + low rating correlations
  "5eaef8702b68455d6e130595_ptsga",
  "5f0f0a2a8b2a480447f31b21_lqgpz",
  "611d03b822d4c8e041ea0c32_m0knb"
)
outliers_partial <- c(
  "5dc3485219ca0326027ce91f_37ho9",
  "5c6414540821d30001046198_x9q7r",
  "60dd7b03f1e72d38230df476_9yh9n",
  "5962799cb752840001ca478b_jh4sl",
  "5f44c23fbf2ddb80bcdf0edc_dnbny",
  "5e80370d48b5f47170e30e5c_5w2gf"
)
```

We removed `r length(outliers)` participants upon inspection of...

### Extreme Items

```{r message=FALSE, warning=FALSE}
extreme_items <- df |>
  group_by(Stimulus, Belief) |>
  summarize(n = n() / length(unique(df$Participant))) |>
  pivot_wider(values_from = "n", names_from = "Belief") |>
  mutate(File = paste0("experiment/stimuli/AMFD/", Stimulus)) |>
  arrange(Real) |>
  filter(Real < 0.15 | Real > 0.85)

p_item <- df |>
  filter(Stimulus %in% extreme_items$Stimulus) |>
  mutate(Stimulus = fct_relevel(Stimulus, as.character(extreme_items$Stimulus))) |>
  ggplot(aes(x = Belief_Answer, y = Stimulus, fill = Stimulus)) +
  ggdist::stat_slab(scale = 2, slab_alpha = 0.9, normalize = "groups") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  ggimage::geom_image(data = extreme_items, aes(image = File, x = 0, y = Stimulus), size = 0.1, by = "height") +
  # scale_y_discrete(expand = c(0.5, 0.5)) +
  scale_x_continuous(
    limits = c(-1, 1),
    expand = c(0, 0),
    breaks = c(-1, 0, 1),
    label = c("Fake", "", "Real")
  ) +
  scale_fill_viridis_d(option = "inferno") +
  labs(x = "Simulation Monitoring", y = "Participants", title = "Distribution of Reality Judgments") +
  guides(fill = "none") +
  see::theme_modern() +
  theme(
    # axis.text.y = element_blank(),
    plot.title = element_text(face = "bold", hjust = 0.5)
  )
# p_item

df <- df |>
  filter(!Stimulus %in% extreme_items$Stimulus)
```

We removed `r nrow(extreme_items)` trials per participant.

### Attention Checks and Duration

```{r message=FALSE, warning=FALSE}
dfsub <- df |>
  group_by(Participant) |>
  select(Participant, starts_with("Attention"), starts_with("Duration"), n_Fake) |>
  slice(1) |>
  ungroup() |>
  rowwise() |>
  mutate(Attention_Check = mean(c(Attention_Check1, Attention_Check2, Attention_Check3))) |>
  ungroup() |>
  arrange(Attention_Check)
```

### Ratings

```{r message=FALSE, warning=FALSE}
dfsub$r_Trustworthy <- NA
dfsub$r_Attractive <- NA
dfsub$r_Beauty <- NA
for (participant in dfsub$Participant) {
  dfsub[dfsub$Participant == participant, "r_Trustworthy"] <- cor(df[df$Participant == participant, "Trustworthy"], df[df$Participant == participant, "Norms_Trustworthy"])
  dfsub[dfsub$Participant == participant, "r_Attractive"] <- cor(df[df$Participant == participant, "Attractive"], df[df$Participant == participant, "Norms_Attractive"])
  dfsub[dfsub$Participant == participant, "r_Beauty"] <- cor(df[df$Participant == participant, "Beauty"], df[df$Participant == participant, "Norms_Attractive"])
}
```

### Summary

```{r message=FALSE, warning=FALSE}
data.frame(Participant = c(paste0("Total (n=", nrow(dfsub), ")")), t(sapply(dfsub[2:ncol(dfsub)], mean, na.rm = TRUE))) |>
  rbind(dfsub) |>
  mutate(Attention_Check = paste0(
    insight::format_value(Attention_Check, 1),
    " (", insight::format_value(Attention_Check1, 1),
    ", ",
    insight::format_value(Attention_Check2, 1),
    ", ",
    insight::format_value(Attention_Check3, 1),
    ")"
  )) |>
  select(-Attention_Check1, -Attention_Check2, -Attention_Check3) |>
  datawizard::data_relocate("Attention_Check", 2) |>
  knitr::kable() |>
  kableExtra::row_spec(1, italic = TRUE) |>
  kableExtra::row_spec(which(dfsub$Participant %in% outliers) + 1, background = "#EF9A9A") |>
  kableExtra::row_spec(which(dfsub$Participant %in% outliers_partial) + 1, background = "#FFCC80")
# kableExtra::row_spec(which(str_detect(dfsub$Participant, "613a972033d79df11a6570de")) + 1, background = "green")
```

```{r message=FALSE, warning=FALSE}
p_att <- dfsub |>
  select(Participant, starts_with("Att")) |>
  pivot_longer(-Participant) |>
  # mutate(name = str_remove(name, "Cor_")) |>
  ggplot(aes(x = Participant, y = value)) +
  geom_bar(aes(fill = name), stat = "identity", position = "dodge") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values = c("black", "#2196F3", "#3F51B5", "#673AB7")) +
  scale_color_manual(values = c("black", "#2196F3", "#3F51B5", "#673AB7")) +
  see::theme_modern() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top"
  ) +
  labs(y = "Score", fill = "") +
  guides(color = "none") +
  ggside::geom_ysidedensity(aes(x = after_stat(scaled), color = name)) +
  ggside::theme_ggside_void() +
  ggside::scale_ysidex_continuous(expand = c(0, 0))

p_time <- dfsub |>
  select(Participant, starts_with("Duration")) |>
  pivot_longer(-Participant) |>
  mutate(name = str_remove(name, "Duration_")) |>
  ggplot(aes(x = Participant, y = value)) +
  geom_bar(aes(fill = name), stat = "identity", position = "dodge") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values = c("#4CAF50", "#FF9800")) +
  scale_color_manual(values = c("#4CAF50", "#FF9800")) +
  see::theme_modern() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top"
  ) +
  labs(y = "Duration (min)", fill = "") +
  guides(color = "none") +
  ggside::geom_ysidedensity(aes(x = after_stat(scaled), color = name)) +
  ggside::theme_ggside_void() +
  ggside::scale_ysidex_continuous(expand = c(0, 0))

p_cor <- dfsub |>
  select(Participant, starts_with("r_")) |>
  pivot_longer(-Participant) |>
  mutate(name = str_remove(name, "r_")) |>
  ggplot(aes(x = Participant, y = value)) +
  geom_bar(aes(fill = name), stat = "identity", position = "dodge") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values = c("#F44336", "#9C27B0", "#5227b0")) +
  scale_color_manual(values = c("#F44336", "#9C27B0", "#5227b0")) +
  see::theme_modern() +
  theme(
    axis.text.x = element_text(
      angle = 45, hjust = 1,
      color = ifelse(levels(dfsub$Participant) %in% outliers, "red", ifelse(levels(dfsub$Participant) %in% outliers_partial, "orange", "black"))
    ),
    legend.position = "top"
  ) +
  labs(y = "Correlation", fill = "") +
  guides(color = "none") +
  ggside::geom_ysidedensity(aes(x = after_stat(scaled), color = name)) +
  ggside::theme_ggside_void() +
  ggside::scale_ysidex_continuous(expand = c(0, 0))

(p_att + theme(axis.text.x = element_blank())) /
  (p_time + theme(axis.text.x = element_blank())) /
  (p_cor)

df <- df |>
  filter(!Participant %in% outliers)
```

## Participants

```{r message=FALSE, warning=FALSE}
dfsub <- df |>
  group_by(Participant) |>
  select(Participant, Age, Sex, Sexual_Orientation, Ethnicity, Education, Nationality, Device_OS, starts_with("Screen"), starts_with("IPIP"), starts_with("Social_"), starts_with("FFNI_"), starts_with("GPTS_"), starts_with("IUS_"), starts_with("SelfAttractiveness"), starts_with("AI"), n_Real, Confidence_Fake, Confidence_Real) |>
  slice(1) |>
  ungroup()
```

The final sample included `r report::report_participants(dfsub, age="Age", sex="Sex")`.

```{r }
plot_distribution <- function(dfsub, what = "Age", title = what, subtitle = "", fill = "orange") {
  dfsub |>
    ggplot(aes_string(x = what)) +
    geom_density(fill = fill) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    ggtitle(title, subtitle = subtitle) +
    theme_modern() +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      plot.subtitle = element_text(face = "italic", hjust = 0.5),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      axis.text.y = element_blank()
    )
}

plot_waffle <- function(dfsub, what = "Nationality", title = what, rows = 8, size = 3) {
  # library(emojifont)
  ggwaffle::waffle_iron(dfsub, what, rows = rows) |>
    # mutate(label = emojifont::fontawesome('fa-smiley')) |>
    # mutate(label = emojifont::emoji('smiley')) |>
    ggplot(aes(x, y)) +
    geom_point(aes(color = group), shape = "square", size = size) +
    # ggwaffle::geom_waffle(color = "white") +
    # geom_point() +
    # geom_text(aes(color=group ,label=label), family='fontawesome-webfont', size=4) +
    # geom_text(aes(color=group ,label=label), family='EmojiOne', size=4) +
    coord_equal() +
    ggtitle(title) +
    labs(fill = "", color = "") +
    # scale_x_continuous(expand = c(0, 0)) +
    # scale_y_continuous(expand = c(0, 0)) +
    theme_void() +
    # ggwaffle::theme_waffle() +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))
}
```

```{r fig.width=20, fig.height=15}
p1 <- estimate_density(dfsub$Age) |>
  ggplot(aes(x = x, y = y)) +
  geom_area(fill = "#FF9800")

p2 <- plot_waffle(dfsub, "Sex") +
  scale_color_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63", "Other" = "#FF9800"))

p3 <- plot_waffle(dfsub, "Sexual_Orientation")

p4 <- plot_waffle(dfsub, "Education") +
  scale_fill_viridis_d()

p5 <- dfsub |>
  group_by(Nationality) |>
  mutate(n = n()) |>
  ungroup() |>
  mutate(Nationality = fct_reorder(Nationality, desc(n))) |>
  ggplot(aes(Nationality)) +
  geom_bar(aes(fill = Nationality)) +
  scale_fill_viridis_d(guide = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p6 <- plot_waffle(dfsub, "Ethnicity") +
  scale_fill_manual(values = c("Latino" = "#FF5722", "Asian" = "#FF9800", "Caucasian" = "#2196F3", "African" = "#4CAF50", "Jewish" = "#9C27B0", "Mixed" = "#795548"))

p7 <- plot_waffle(dfsub, "Screen_Resolution", title = "Screen Resolution") +
  scale_fill_pizza_d() +
  guides(fill = "none")

p8 <- plot_waffle(dfsub, "Device_OS", title = "Device OS") +
  scale_fill_bluebrown_d()

# p10 <- plot_waffle(dfsub, "Screen_Refresh") +
#   scale_fill_viridis_d()

patchwork::wrap_plots(list(p1, p2, p3, p4, p5, p6, p8))
```

# Results

## Manipulation Check

### Real / Fake

```{r message=FALSE, warning=FALSE, fig.height=fig.width*1.5}
# plot(estimate_density(filter(df, Participant == "60dd7b03f1e72d38230df476_9yh9n")$Belief_Answer))
df |>
  mutate(Participant = fct_relevel(Participant, df |>
    group_by(Participant) |>
    summarize(Belief_Answer = mean(Belief_Answer)) |>
    ungroup() |>
    arrange(Belief_Answer) |>
    pull(Participant) |>
    as.character())) |>
  # mutate(Participant = fct_relevel(Participant, as.character(dfsub$Participant))) |>
  ggplot(aes(x = Belief_Answer, y = Participant, fill = Participant)) +
  ggdist::stat_slab(scale = 2, slab_alpha = 0.9, normalize = "groups", color = "black", size = 0.1) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  scale_y_discrete(expand = c(0.02, 0)) +
  scale_x_continuous(
    limits = c(-1, 1),
    expand = c(0, 0),
    breaks = c(-0.95, 0, 0.95),
    label = c("Fake", "", "Real")
  ) +
  scale_fill_viridis_d() +
  labs(x = "Simulation Monitoring", y = "Participants", title = "Distribution of Reality Judgments") +
  guides(fill = "none") +
  see::theme_modern() +
  theme(
    axis.text.y = element_blank(),
    plot.title = element_text(face = "bold", hjust = 0.5)
  ) +
  ggside::geom_xsidedensity(fill = "grey", color = "white") +
  ggside::scale_xsidey_continuous(expand = c(0, 0))
```

### Colinearity

```{r message=FALSE, warning=FALSE}
IVs <- c("Attractive", "Beauty", "Trustworthy", "Familiar")

preds <- data.frame()
dats <- data.frame()
for (x in IVs) {
  for (y in IVs) {
    if (x == y) next
    print(paste(y, "~", x))
    model <- glmmTMB::glmmTMB(as.formula(
      paste(y, "~", x, "* Sex * Stimulus_Interest + (1|Participant) + (1|Stimulus)")
    ),
    data = df,
    family = glmmTMB::beta_family()
    )

    # model <- mgcv::gamm(Real ~ s(Attractive) + Trustworthy,
    #                     random = list(Participant=~1, Stimulus=~1),
    #                     data = df,
    #                     family=mgcv::betar())

    pred <- estimate_relation(model, at = c(x, "Stimulus_Interest", "Sex"), length = 20)
    pred$y <- y
    pred <- data_rename(pred, x, "Score")
    pred$x <- x
    preds <- rbind(preds, pred)

    dats <- rbind(dats, data.frame(Score = df[[x]], Predicted = df[[y]], x = x, y = y, Sex = df$Sex))
  }
}

dats <- mutate(dats, x = fct_relevel(x, IVs), y = fct_relevel(y, IVs))
preds <- mutate(preds, x = fct_relevel(x, IVs), y = fct_relevel(y, IVs))
```

```{r message=FALSE, warning=FALSE}
dats |>
  ggplot(aes(x = Score, y = Predicted)) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = FALSE) +
  # geom_ribbon(data = preds, aes(ymin = CI_low, ymax = CI_high, group = Stimulus_SameSex), alpha = 0.3) +
  geom_line(data = preds, aes(color = Sex, linetype = Stimulus_Interest)) +
  scale_fill_gradientn(colors = c("white", "#FF9800", "#F44336"), guide = "none") +
  scale_color_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63")) +
  scale_linetype_manual(values = c("TRUE" = "solid", "FALSE" = "dashed")) +
  scale_x_continuous(expand = c(0, 0), labels = scales::percent()) +
  scale_y_continuous(expand = c(0, 0), labels = scales::percent()) +
  facet_grid(y ~ x, switch = "both") +
  theme_modern() +
  labs(title = "Collinearity in the Stimuli Ratings") +
  theme(
    aspect.ratio = 1,
    strip.background = element_blank(),
    strip.placement = "outside",
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(face = "bold", hjust = 0.5)
  ) +
  ggnewscale::new_scale_fill() +
  scale_fill_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63")) +
  ggside::geom_xsidedensity(aes(fill = Sex), color = NA, alpha = 0.3) +
  ggside::geom_ysidedensity(aes(fill = Sex), color = NA, alpha = 0.3) +
  ggside::theme_ggside_void() +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::ggside(collapse = "all")
```

### Effect of Delay

```{r message=FALSE, warning=FALSE}
model <- glmmTMB::glmmTMB(Belief ~ Delay + (1 | Participant) + (1 | Stimulus),
  data = df,
  family = "binomial"
)
pred <- estimate_relation(model, at = "Delay", length = 20)

m_conf <- glmmTMB::glmmTMB(Belief_Confidence ~ Belief / Delay + ((Belief / Delay) | Participant) + (1 | Stimulus),
  data = df,
  family = glmmTMB::beta_family()
)
y_conf <- estimate_relation(m_conf, at = c("Delay", "Belief"), length = 20)
y_conf <- y_conf |>
  mutate_at(c("Predicted", "CI_low", "CI_high"), function(x) ifelse(y_conf$Belief == "Real", datawizard::rescale(x, range = c(0, 1), to = c(0.5, 1)), datawizard::rescale(x, range = c(1, 0), to = c(0, 0.5))))

df |>
  ggplot(aes(x = Delay, y = Real)) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = FALSE) +
  geom_hline(yintercept = 0.5, linetype = "dotted") +
  # geom_ribbon(data=y_conf, aes(y=Predicted, ymin = CI_low, ymax = CI_high), alpha = 0.3) +
  geom_line(data = y_conf, aes(y = Predicted, group = Belief), linetype = "dashed", color = "red") +
  geom_ribbon(data = pred, aes(y = Predicted, ymin = CI_low, ymax = CI_high), alpha = 0.3) +
  geom_line(data = pred, aes(y = Predicted), color = "red") +
  scale_fill_gradientn(colors = c("white", "#795548"), guide = "none") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0), labels = scales::percent) +
  theme_modern() +
  labs(title = "Effect of Re-exposure Delay", x = "Minutes") +
  theme(
    aspect.ratio = 1,
    strip.background = element_blank(),
    strip.placement = "outside",
    plot.title = element_text(face = "bold", hjust = 0.5)
  ) +
  ggside::geom_xsidedensity(fill = "#795548", color = "white") +
  ggside::geom_ysidedensity(fill = "#9C27B0", color = "white") +
  ggside::theme_ggside_void() +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::ggside(collapse = "all")

parameters::parameters(model)
```

## Determinants of Reality

```{r message=FALSE, warning=FALSE}
make_model <- function(df, var = "Attractive", formula = var, fill = "#2196F3") {
  # Models
  m_real <- glmmTMB::glmmTMB(as.formula(paste0("Belief ~ ", formula)),
    data = df,
    family = "binomial"
  )
  y_real <- estimate_relation(m_real, at = c(var, "Sex"), length = 21)
  # gam <- brms::brm(paste0("Belief ~ s(", var, ", by=Sex) + (1|Participant) + (1|Stimulus)"),
  #                  data=df,
  #                  algorithm="sampling",
  #                  family = "bernoulli")
  # trend <- estimate_relation(gam, at = c(var, "Sex"), length = 81, preserve_range=FALSE)
  # slope <- estimate_slopes(gam, trend=var, at = c(var, "Sex"), length = 81)
  # trend$Trend <- interpret_pd(slope$pd)
  # trend$group <- 0
  # trend$group[2:nrow(trend)] <- as.character(cumsum(ifelse(trend$Trend[2:nrow(trend)] == trend$Trend[1:nrow(trend)-1], 0, 1)))

  m_conf <- glmmTMB::glmmTMB(as.formula(paste0("Belief_Confidence ~ Belief /", formula)),
    data = df,
    family = glmmTMB::beta_family()
  )
  y_conf <- estimate_relation(m_conf, at = c(var, "Belief", "Sex"), length = 21)
  y_conf <- y_conf |>
    mutate_at(c("Predicted", "CI_low", "CI_high"), function(x) ifelse(y_conf$Belief == "Real", datawizard::rescale(x, range = c(0, 1), to = c(0.5, 1)), datawizard::rescale(x, range = c(1, 0), to = c(0, 0.5))))

  # Significance
  param <- parameters::parameters(m_real, effects = "fixed", keep = var)
  sig <- data.frame(
    x = 0.5,
    y = y_real[c(11, 32), "Predicted"] + c(0.05, -0.05),
    p = c(min(param$p[c(1, 3)]), min(param$p[c(2, 4)])),
    Sex = c("Female", "Male")
  )
  sig$label <- format_p(sig$p, stars_only = TRUE)

  # Plot
  p <- df |>
    ggplot(aes_string(x = var, y = "Real")) +
    stat_density_2d(aes(fill = ..density..), geom = "raster", contour = FALSE) +
    scale_fill_gradientn(colors = c("white", fill), guide = "none") +
    ggnewscale::new_scale_fill() +
    geom_hline(yintercept = 0.5, linetype = "dotted") +
    # geom_point2(alpha = 0.25, size = 4, color = "black") +
    geom_line(data = y_conf, aes(y = Predicted, group = interaction(Belief, Sex), color = Sex), linetype = "dashed") +
    geom_ribbon(data = y_real, aes(y = Predicted, group = Sex, fill = Sex, ymin = CI_low, ymax = CI_high), alpha = 1 / 3) +
    geom_line(data = y_real, aes(y = Predicted, color = Sex), size=1) +
    # geom_ribbon(data = trend, aes(y = Predicted, group=Sex, fill=Sex, ymin = CI_low, ymax = CI_high), alpha = 1/6) +
    # geom_line(data = trend, aes(y = Predicted, color=Sex, linetype=Trend, group=interaction(Sex, group)), size=0.6) +
    geom_text(data = sig, aes(y = y, x = x, label = label, color = Sex), size = 8) +
    scale_color_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63")) +
    scale_fill_manual(values = c("Male" = "#2196F3", "Female" = "#E91E63")) +
    scale_x_continuous(expand = c(0, 0), labels = scales::percent) +
    scale_y_continuous(expand = c(0, 0), labels = scales::percent) +
    guides(fill = guide_legend(override.aes = list(alpha = 1))) +
    theme_modern(axis.title.space = 5) +
    ggside::geom_xsidedensity(fill = fill, color = "white") +
    ggside::geom_ysidedensity(fill = "#9C27B0", color = "white") +
    ggside::theme_ggside_void() +
    ggside::scale_ysidex_continuous(expand = c(0, 0)) +
    ggside::scale_xsidey_continuous(expand = c(0, 0))

  list(p = p, model_belief = m_real, model_confidence = m_conf)
}
```

```{r message=FALSE, warning=FALSE}
rez_at <- make_model(filter(df, Stimulus_Interest == TRUE),
  formula = "Sex / poly(Attractive, 2) + (1|Participant) + (1|Stimulus)",
  var = "Attractive", fill = "#F44336"
)
rez_gl <- make_model(filter(df, Stimulus_Interest == TRUE),
  formula = "Sex / poly(Beauty, 2) + Trustworthy + (1|Participant) + (1|Stimulus)",
  var = "Beauty", fill = "#E91E63"
)
rez_tr <- make_model(filter(df, Stimulus_Interest == TRUE),
  formula = "Sex / poly(Trustworthy, 2) + Beauty + (1|Participant) + (1|Stimulus)",
  var = "Trustworthy", fill = "#4CAF50"
)
rez_fa <- make_model(filter(df, Stimulus_Interest == TRUE),
  formula = "Sex / poly(Familiar, 2) + (1|Participant) + (1|Stimulus)",
  var = "Familiar", fill = "#2196F3"
)

(rez_at$p +
  theme(axis.text.x = element_blank()) +
  labs(x = "Attractiveness") |
  rez_gl$p +
    labs(x = "Beauty") +
    theme(
      axis.text.x = element_blank(),
      axis.title.y = element_blank(),
      axis.text.y = element_blank()
    )
) /
  (rez_tr$p +
    labs(x = "Trustworthiness") |
    rez_fa$p +
      labs(x = "Familiarity") +
      theme(
        axis.text.y = element_blank(),
        axis.title.y = element_blank()
      )
  ) +
  plot_annotation(title = "Determinants of Reality Beliefs", theme = theme(plot.title = element_text(face = "bold", hjust = 0.5))) +
  plot_layout(guides = "collect")
```

```{r message=FALSE, warning=FALSE}
parameters::parameters(rez_at$model_belief, effects = "fixed", keep = "Attractive")
performance::r2(rez_at$model_belief)
performance::icc(rez_at$model_belief, by_group = TRUE)
parameters::parameters(rez_at$model_confidence, effects = "fixed", keep = "Attractive")

parameters::parameters(rez_gl$model_belief, effects = "fixed", keep = "Beauty")
performance::r2(rez_gl$model_belief)
performance::icc(rez_gl$model_belief, by_group = TRUE)
parameters::parameters(rez_gl$model_confidence, effects = "fixed", keep = "Beauty")

parameters::parameters(rez_tr$model_belief, effects = "fixed", keep = "Trustworthy")
performance::r2(rez_tr$model_belief)
performance::icc(rez_tr$model_belief, by_group = TRUE)
parameters::parameters(rez_tr$model_confidence, effects = "fixed", keep = "Trustworthy")

parameters::parameters(rez_fa$model_belief, effects = "fixed", keep = "Familiar")
performance::r2(rez_fa$model_belief)
performance::icc(rez_fa$model_belief, by_group = TRUE)
parameters::parameters(rez_fa$model_confidence, effects = "fixed", keep = "Familiar")
```

## Interaction

```{r message=FALSE, warning=FALSE}
cor_test(dfsub, "SelfAttractiveness1", "SelfAttractiveness2")

df$Self_Attractiveness <- rowMeans(df[c("SelfAttractiveness1", "SelfAttractiveness2")])
```

```{r message=FALSE, warning=FALSE}
m_real <- glmmTMB::glmmTMB(Belief ~ Sex / (poly(Attractive, 2) * Self_Attractiveness) + (1 | Participant) + (1 | Stimulus),
  data = filter(df, Stimulus_Interest == TRUE),
  family = "binomial"
)

parameters::parameters(m_real, effects = "fixed", keep = "Self_Attractiveness")

m_real <- glmmTMB::glmmTMB(Belief ~ Sex / (poly(Beauty, 2) * Self_Attractiveness) + Trustworthy + (1 | Participant) + (1 | Stimulus),
  data = filter(df, Stimulus_Interest == TRUE),
  family = "binomial"
)

parameters::parameters(m_real, effects = "fixed", keep = "Self_Attractiveness")
```

## Inter-Individual Correlates

```{r message=FALSE, warning=FALSE}
model <- glmmTMB::glmmTMB(Belief ~ 1 + (1 | Participant) + (1 | Stimulus), data = df, family = "binomial")
dfsub <- modelbased::estimate_grouplevel(model) |>
  filter(Group == "Participant") |>
  transmute(Participant = Level, Simulation_Monitoring = Coefficient) |>
  merge(dfsub, by = "Participant")

model <- glmmTMB::glmmTMB(Belief_Confidence ~ Belief + (Belief | Participant) + (1 | Stimulus), data = df, family = glmmTMB::beta_family())
dfsub <- modelbased::estimate_grouplevel(model) |>
  filter(Group == "Participant") |>
  transmute(Participant = Level, Parameter, Coefficient) |>
  pivot_wider(names_from = "Parameter", values_from = "Coefficient") |>
  transmute(Participant, Confidence_Real_Model = BeliefReal, Confidence_Fake_Model = `(Intercept)`) |>
  merge(dfsub, by = "Participant")
```
```{r message=FALSE, warning=FALSE}
correlation::correlation(
  dfsub[c("Simulation_Monitoring", "Confidence_Fake_Model", "Confidence_Real_Model")],
  dfsub[c("Confidence_Fake", "Confidence_Real", "n_Real")],
)
```

### IPIP-6

```{r message=FALSE, warning=FALSE}
make_correlation <- function(x, y) {
  cor <- correlation::correlation(x,
    y,
    bayesian = TRUE,
    bayesian_prior = "medium.narrow",
    sort = TRUE
  ) |>
    datawizard::data_remove(c("ROPE_Percentage"))
  cor$`BF (Spearman)` <- format_bf(
    correlation::correlation(
      x, y,
      bayesian = TRUE,
      ranktransform = TRUE,
      bayesian_prior = "medium.narrow"
    )$BF,
    name = NULL, stars = TRUE
  )
  cor |>
    arrange(desc(BF))
}

plot_correlation <- function(dfsub, x = "Confidence_Real", y = "IPIP6_Openness", xlab = x, ylab = y, fill = "grey") {
  param <- cor_test(dfsub, x, y, bayesian = TRUE)

  # Format stat output
  r <- str_replace(str_remove(insight::format_value(param$rho), "^0+"), "^-0+", "-")
  CI_low <- str_replace(str_remove(insight::format_value(param$CI_low), "^0+"), "^-0+", "-")
  CI_high <- str_replace(str_remove(insight::format_value(param$CI_high), "^0+"), "^-0+", "-")

  stat <- paste0("italic(r)~'= ", r, ", 95% CI [", CI_low, ", ", CI_high, "], BF'['10']~'", paste0(insight::format_bf(param$BF, name = "")), "'")

  label <- data.frame(
    x = min(dfsub[[x]], na.rm = TRUE),
    y = max(dfsub[[y]], na.rm = TRUE),
    label = stat
  )

  # Plot
  dfsub |>
    ggplot(aes_string(x = x, y = y)) +
    geom_point2(
      size = 3,
      color = fill,
      # color = DVs[x],
      alpha = 2 / 3
    ) +
    geom_smooth(method = "lm", color = "black", formula = "y ~ x", alpha = 0.3) +
    labs(y = ylab, x = xlab) +
    geom_label(data = label, aes(x = x, y = y), label = str2expression(label$label), hjust = 0, vjust = 1) +
    theme_modern() +
    ggside::geom_xsidedensity(fill = "grey", color = "white") +
    ggside::geom_ysidedensity(fill = fill, color = "white") +
    ggside::theme_ggside_void() +
    ggside::scale_ysidex_continuous(expand = c(0, 0)) +
    ggside::scale_xsidey_continuous(expand = c(0, 0))
}
```
```{r message=FALSE, warning=FALSE}
sr <- c("Simulation_Monitoring", "Confidence_Fake_Model", "Confidence_Real_Model", "Confidence_Fake", "Confidence_Real", "n_Real")

r <- make_correlation(dfsub[sr], select(dfsub, starts_with("IPIP")))
filter(r, BF > 1)

plot_correlation(dfsub,
  x = "Confidence_Real",
  y = "IPIP6_HonestyHumility",
  xlab = "Confidence_Real",
  ylab = "IPIP6_HonestyHumility",
  fill = "grey"
)
```

### AI

```{r message=FALSE, warning=FALSE}
rez <- parameters::n_factors(select(dfsub, starts_with("AI")))
plot(rez)

efa <- parameters::factor_analysis(select(dfsub, starts_with("AI")), n = 3, rotation = "varimax", sort = TRUE)
efa
dfsub <- predict(efa, names = c("AI_Enthusiasm", "AI_Realness", "AI_Danger")) |>
  cbind(dfsub)
```
```{r message=FALSE, warning=FALSE}
r <- make_correlation(dfsub[sr], select(dfsub, AI_Enthusiasm, AI_Realness, AI_Danger))
filter(r, BF > 1)
```

### Social Anxiety

```{r message=FALSE, warning=FALSE}
r <- make_correlation(dfsub[sr], select(dfsub, starts_with("Social_")))
filter(r, BF > 1)
```

### Intolerance to Uncertainty

```{r message=FALSE, warning=FALSE}
r <- make_correlation(dfsub[sr], select(dfsub, starts_with("IUS_")))
filter(r, BF > 1)
```

### Narcissism

```{r message=FALSE, warning=FALSE}
r <- make_correlation(dfsub[sr], select(dfsub, starts_with("FFNI_")))
filter(r, BF > 1)
```

### Paranoid Beliefs

```{r message=FALSE, warning=FALSE}
r <- make_correlation(dfsub[sr], select(dfsub, starts_with("GPTS_")))
filter(r, BF > 1)
```