### 2024-10-08 submission/new_manuscript/manuscript.qmd
changes to manuscript and R1

---
title: "Too Beautiful to be Fake: Attractive Faces are Less Likely to be Judged as Artificially Generated"
shorttitle: "Attractiveness and Reality"
author:
  - name: Dominique Makowski
    corresponding: true
    orcid: 0000-0001-5375-9967
    email: D.Makowski@sussex.ac.uk
    url: https://realitybending.github.io/
    roles:
      - Conceptualization
      - Data curation
      - Formal Analysis
      - Funding acquisition
      - Investigation
      - Methodology
      - Project administration
      - Resources
      - Software
      - Supervision
      - Validation
      - Visualization
      - Writing – original draft
    affiliations:
      - name: University of Sussex
        department: School of Psychology
  - name: An Shu Te
    #corresponding: 
    orcid: 0000-0002-9312-5552
    role:
      - Project administration
      - Resources
      - Writing – original draft
    affiliations:
      - name: Nanyang Technological University
        department: School of Social Sciences
  - name: Stephanie Kirk  # stephanie.morgank@gmail.com
    orcid: 0000-0002-9312-5552
    role:
      - Project administration
      - Resources
      - Writing – original draft
    affiliations:
      - name: Nanyang Technological University
        department: School of Social Sciences
  - name: Ngoi Zi Liang  # ziliang.ngoi@ntu.edu.sg
    #affiliation   : "2"
    role:
      - Project administration
      - Resources
      - Writing – review & editing
    affiliations:
      - name: Nanyang Technological University
        department: School of Social Sciences
  - name: Panagiotis Mavros  # panos.mavros@telecom-paris.fr
    #affiliation   : "3"
    orcid: 0000-0002-1540-5516
    role:
      - Supervision
      - Writing – review & editing
    affiliations:
      - name:  Télécom Paris
        department: Department of Economics and Social Sciences
  - name: S.H. Annabel Chen  # AnnabelChen@ntu.edu.sg
    role:
      - Project administration
      - Supervision
      - Writing – review & editing
    affiliations:
      - name: University of Sussex
        department: School of Psychology
      - name: Nanyang Technological University
        department: LKC Medicine
      - name: National Institute of Education
      - name: Nanyang Technological University
        department:  Centre for Research and Development in Learning
author-note:
  status-changes: 
    # Example: [Author name] is now at [affiliation].
    affiliation-change: null
    # Example: [Author name] is deceased.
    deceased: null
  # Disclosures condensed to one paragraph, but you can start a field with two line breaks to break them up: \n\nNew Paragraph
  disclosures:
    # Acknowledge and cite data/materials to be shared.
    # Example. Because the authors are equal contributors, order of authorship was determined by a fair coin toss.
    # Example: This study was registered at X (Identifier Y).
    study-registration: null
    data-sharing: null
    # Example: This article is based on the dissertation completed by [citation].  
    related-report: null
    # Example: [Author name] has been a paid consultant for Corporation X, which funded this study.
    conflict-of-interest: null
    # Example: This study was supported by Grant [Grant Number] from [Funding Source].
    financial-support: null
    # Example: The authors are grateful to [Person] for [Reason].
    gratitude: null
    authorship-agreements: null
abstract: " Technological advances render the distinction between artificial (e.g., computer-generated faces) and real stimuli increasingly difficult, yet the factors driving our beliefs regarding the nature of ambiguous stimuli remain largely unknown. In this study, 150 participants rated 109 pictures of faces on 4 characteristics (attractiveness, beauty, trustworthiness, familiarity). The stimuli were then presented again with the new information that some of them were AI-generated, and participants had to rate each image according to whether they believed them to be real or fake. Despite all images being pictures of real faces from the same database, most participants did indeed rate a large portion of them as 'fake' (often with high confidence), with strong intra- and inter-individual variability. Our results suggest a gender-dependent role of attractiveness on reality judgements, with faces rated as more attractive being classified as more real. We also report links between reality beliefs tendencies and dispositional traits such as narcissism and paranoid ideation."
keywords: [attractiveness, AI-generated images, fiction, fake news, sense of reality]
floatsintext: true
numbered-lines: true
bibliography: references.bib
suppress-title-page: false
mask: false
# Language options. See https://quarto.org/docs/authoring/language.html
lang: en
language:
  citation-last-author-separator: "and"
  citation-masked-author: "Masked Citation"
  citation-masked-date: "n.d."
  citation-masked-title: "Masked Title"
  title-block-author-note: "Author Note"
  title-block-correspondence-note: "Correspondence concerning this article should be addressed to"
  title-block-role-introduction: "Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows:"
  references-meta-analysis:
format:
  apaquarto-docx: default
  apaquarto-html: default
  apaquarto-pdf:
    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
    documentmode: man
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| label: setup
#| include: false
library(conflicted)
library(tidyverse)
library(flextable)
library(ftExtra)
library(officer)
library(knitr)

conflicts_prefer(dplyr::filter, .quiet = TRUE)
conflicts_prefer(flextable::separate_header, .quiet = TRUE)
```

```{r, echo = FALSE, warning=FALSE, message=FALSE}
# options and parameters
options(digits = 3)

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.width = see::golden_ratio(9),
  fig.height = 9,
  fig.path = "figures/"
)

cache <- TRUE
```

<!-- stakes and Post-truth era -->

Advancements in technology have now made it possible to create
near-perfect simulations that are indistinguishable from reality. These
simulations are more affordable, faster to produce, and more accessible
than at any other time in human history. These artificial, yet realistic
constructs permeate all areas of life through immersive works of
fiction, deep fakes (real-like images and videos generated by deep
learning algorithms), virtual and augmented reality (VR and AR),
artificial beings (artificial intelligence "bots" with or without a
physical form), fake news and skewed narratives, of which ground truth
is often hard to access [@nightingale2022]. Such developments not only
carry important consequences for the technological and entertainment
sectors, but also for security and politics - for instance if used for
propaganda and disinformation, recruitment into malevolent
organizations, or religious indoctrination [@pantserev2020]. This issue
is central to what has been coined the "post-truth era"
[@lewandowsky2017beyond], in which the distinction (and lack thereof)
between authentic and simulated objects will play a critical role.

<!-- Ambiguous stimuli and irrelevance of objective features -->

While not all simulations have achieved perfect realism, such as
Computer Generated Images (CGI) in movies or via recent algorithms such
as GANs or diffusion model, which often include distortions or lack
certain key details distinguishing them from real images
[@mcdonnell2010face; @corvi2022detection], it is fair to assume that
these technical limitations will become negligible in the near future.
This is particularly true in the field of face generation, where
face-generation algorithms are already able to create stimuli that are
virtually indistinguishable from real photos [@nightingale2022;
@tucciarelli2020; @moshel2022]. Such a technological feat, however,
leads to a new question: if real and fake stimuli cannot be
differentiated based on their objective "physical" characteristics, how
can we form judgements regarding their nature?

<!-- Context -->

Literature shows that the context surrounding a stimulus often plays an
important role in the assessment of its reality [a process henceforth
referred to as *simulation monitoring*, @makowski2018cognitive;
@makowski2019phenomenal]. With the extensive search and processing of
cues within ambiguous stimuli being an increasingly complex and
cognitively effortful strategy [@michael2021source;
@susmann2021persuasion], people tend to draw on peripheral contextual
cues (**Figure 1**), such as the source of the stimulus (e.g., which
journal was the information published in), and its credibility,
authority and expertise, to help facilitate their evaluation
[@petty1986elaboration; @susmann2021persuasion; @michael2021source].
However, the atomization and decontextualization of information allowed
by online social media (where text snippets or video excerpts are often
mass-shared with little context) makes this task progressively difficult
[@berghel2018weaponizing; @chen2015news]. Thus, in the absence of clear
contextual information, what drives our beliefs of reality?

```{r message=FALSE, warning=FALSE, fig.cap="Figure 1. The decision to believe that an ambiguous stimulus (of any form, e.g., images, text, videos, environments, ...) is real or fake depends of individual characteristics (e.g., personality and cognitive styles), stimulus-related features (context, emotionality), and their interaction, which can manifest for instance in our bodily reaction.", out.width="100%"}
knitr::include_graphics("../../figures/figure1.png")
```

<!-- Internal features -->

Evidence suggests that inter-individual characteristics play a crucial
role in simulation monitoring, with factors such as cognitive style,
prior beliefs, and personality traits [@bryanov2021determinants;
@ecker2022psychological; @sindermann2020short]. For instance,
individuals with stronger analytical reasoning skills have been found to
better discriminate real from fake stimuli [@pehlivanoglu2021role;
@pennycook2019lazy], and prior knowledge or beliefs about the stimulus
influences one's perception of it by biasing the attention deployment
towards information that is in line with one's expectations
[@britt2019reasoned]. Furthermore, dispositional traits, such as high
levels of narcissism and low levels of openness and conscientiousness,
have been associated with greater susceptibility to fake news
[@piksa2022cognitive; @sindermann2020short]. **Interestingly, a recent
review suggested that narcissism was related to a strong self-perceived
ability at detecting lies [@turi2022tangled], which could translate to
participants scoring high on narcissism providing more clear cut and
confidence responses. Conversely, those high in honesty-humility tend to
be more conservative in their judgments to ensure fairness
[@liu2020honesty], likely resulting in lower confidence ratings.**

<!-- Reactive features -->

Beyond stimulus- and individual-related characteristics, evidence
suggests that the interaction between the two (i.e., the subjective
reaction associated with the experience of a given stimulus),
contributes to simulation monitoring decisions. For instance, the
intensity of experienced emotions have been shown to increase one's
sense of presence - the extent to which one feels like "being there", as
if the object of experience was real - when engaged in a fictional movie
or a VR environment [@sanchez2005presence; @makowski2017being]. Indeed,
participants' self-reported emotional arousal were found to
significantly predict the probability that they would perceive images as
real [@azevedo2020body]. Conversely, beliefs that emotional stimuli were
fake (e.g., that emotional scenes were not authentic but instead
involved actors and movie makeup) were found to result in emotion
down-regulation [@makowski2019phenomenal; @sperduti2017distinctive]. In
line with these findings, studies on susceptibility to fake news have
also found heightened stimulus emotionality to be associated with
greater belief [@martel2020reliance; @bago2022emotion], and higher
neurophysiological arousal was predictive of judging realistic images as
real [@azevedo2020body]. Additionally, other factors, such as the
stimuli's perceived self-relevance [@sperduti2016paradox;
@goldstein2009pleasure], as well as familiarity [@begg1992dissociation],
could also play a role in guiding our appraisal of a stimulus.**For
instance,** @miller2023 **reported that participants were more likely to
mistakenly identify AI-generated faces as real because they perceived
them as more familiar.**

<!-- Attractiveness and trustworthiness  -->

Due to their popularity as a target of CGI technology and the prospect
offered with facial features that can be experimentally manipulated,
AI-generated images of faces are increasingly used to study face
processing [@dawel2021systematic], in particular in relationship with
saliency or emotions, as well as to other important components of face
evaluation, such as trustworthiness or attractiveness
[@sobieraj2014beautiful; @balas2017; @tsikandilakis2019beauty;
@calbi2017]. Interestingly, artificially created faces rated as more
attractive (by an independent group of raters) were perceived as less
real [@tucciarelli2020]. Conversely, @liefooghe2022faces reports that
attractiveness ratings were significantly lower when participants who
were told that the faces were AI-generated were compared to those who
had no prior knowledge. **Similarly, when participants are informed that
faces are AI-generated, the perceived artificiality leads to lower trust
ratings [@wang2024], even when they are real faces
[@liefooghe2022faces]. In contrast, when participants are unaware that
the faces are AI-generated, trust ratings for these synthetic faces tend
to increase [@nightingale2022]**. Whereas this line of evidence suggests
that reality beliefs have an effect on face attractiveness and
**trustworthiness** ratings, the opposite question - whether
attractiveness **and trustworthiness contribute** to the formation of
reality beliefs - has received little attention to date.

<!-- Goals and hypotheses -->

AI-generated content, in particular realistic images, is becoming
commonplace and carries important risks for misinformation and
black-mailing [@viola2023designed], emphasizing the need to understand
the different components that come into play in the formation of reality
beliefs. This exploratory study primarily aims at investigating the
effect of facial attractiveness on simulation monitoring, i.e., on the
beliefs that an image is real or artificially generated. Based on the
*affective reality theory* [@makowski2023affective], which suggests that
salient and emotional stimuli are perceived to be more real (up to a
point of reversal after which beliefs of fiction becomes used an emotion
regulation strategy), we hypothesize a quadratic relationship between
perceived realness and attractiveness: faces rated as highly attractive
or unattractive will more likely be believed to be real. We expect a
similar relationship with trustworthiness ratings given its
well-established link with attractiveness [@liefooghe2022faces;
@bartosik2021you; @garrido2017kdef; @little2011facial], and a positive
relationship with familiarity (as more familiar faces would appear as
more salient, self-relevant and anchored in reality). Additionally, we
will further explore the link shared by dispositional traits, such as
personality and attitude towards AI, with simulation monitoring
tendencies. Importantly, this study does investigate the discriminative
accuracy between "true" photos and "true" artificially-generated images
(which we consider more a technological issue than a psychological one),
focusing on the beliefs that a stimulus is real or fake, independently
of its true nature.

## Methods

All the material (preregistration, experiment demo, experiment code, raw
data, analysis script with complementary figures and analyses, etc.) is
available at
[**https://github.com/RealityBending/FakeFace**](https://github.com/RealityBending/FakeFace){.uri}.

**Ethics Statement.** This study was approved by the NTU Institutional
Review Board (NTU IRB-2022-187) <!-- [**IRB NUMBER**] --> and all
procedures performed were in accordance with the ethical standards of
the institutional board and with the 1964 Helsinki Declaration. All
participants provided their informed consent prior to participation and
were incentivized after completing the study.

**Procedure.** In the first part of the study, participants answered a
series of personality questionnaires presented in the order below. These
include the *Mini-IPIP6* <!-- (24 items) [@sibley2011]  --> [24 items,
@sibley2011] measuring 6 personality traits, the *SIAS-6* and the
*SPS-6* <!-- (6 items each) [@peters2012development] --> [6 items each,
@peters2012development] assessing social anxiety levels, 5 items we
devised pertaining to expectations about AI-generated image technology
("I think current Artificial Intelligence algorithms can generate very
realistic images"), **to potentially test and mitigate the potential
effect of expectations/beliefs about AI.** **These items were** mixed
with 5 items from the general attitudes towards AI scale to lower the
former's saliency and the possibility of it priming the subjects about
the task, <!-- (*GAAIS*) [@schepman2020initial]. --> [*GAAIS*,
@schepman2020initial] the *FFNI-BF*
<!-- (30 items) [@jauk2022validation] --> [30 items,
@jauk2022validation] measuring 9 facets of narcissism; the *R-GPTS*
<!-- (18 items) [@freeman2021revised] --> [18 items,
@freeman2021revised] measuring 2 dimensions related to paranoid
thinking; and the *IUS-12* <!-- (12 items) [@carleton2007fearing] -->
[12 items, @carleton2007fearing] measuring intolerance to uncertainty.
Self-rated attractiveness was also assessed using 2 items - one
measuring general attractiveness
<!-- ("How attractive would you say you are?") [@marcinkowska2021self] -->
["How attractive would you say you are?", @marcinkowska2021self] and the
other measuring physical attractiveness
<!-- ("How would you rate your own physical attractiveness relative to the average") [@spielmann2020predictive] -->
["How would you rate your own physical attractiveness relative to the
average", @spielmann2020predictive]. 3 attention check questions were
also embedded in the surveys.

In the second part of this study, images of neutral-expression faces
from the validated American Multiracial Face Database
<!-- (AMFD) [@chen2021broadening]  --> [AMFD, @chen2021broadening] were
presented to the participants for 500ms each, in a randomized order,
following a fixation cross display (750 ms). T**he decision to present
the faces for 500 ms was based on pilot studies, which demonstrated that
this duration provides a sufficient perceptual window for
decision-making and aligns with previous research indicating stable
judgment levels and increased confidence beyond this exposure time
[@willis2006first].**

**The AMFD is a recently validated database including a set of 110
pictures of homogeneous quality featuring diverse faces (particularly in
terms of ethnicity), each (except one) posing with either a neutral or
smiling expression. We selected all 109 neutral images (89 women and 20
men) to reduce the influence of confounding factors like affect. The
AMFD primarily features racially ambiguous faces, representing multiple
racial categories such as multiracial, Latinx, and white. The database
includes 81 faces from individuals self-reporting two racial backgrounds
and 29 from those with three or more racial backgrounds: 33%
Asian/White, 22% Latinx/White, 11% Asian/Latinx, 6% White/Middle
Eastern, 5% Black/White, and 5% Asian/Middle Eastern, with about 18%
identifying as other racial backgrounds.**

After each stimulus presentation, ratings of *Trustworthiness* ("I find
this person trustworthy") and *Familiarity* ("This person reminds me of
someone I know") were collected using visual analog scales. Notably, as
facial attractiveness is a multidimensional construct, encompassing
evolutionary, sociocultural, biological as well as cognitive aspects
[@rhodes2006evolutionary; @han2018beauty], we assessed attractiveness
using 2 visual analog scales, measuring general *Attractiveness* ("I
find this person attractive") and physical *Beauty* ("This face is
good-looking"). **This dual-scale approach aims to reflect two
conceptually distinct dimensions: Attractiveness might capture personal,
Self-relevant and subjective appeal, whereas Beauty might be related to
a more “objective” decision based on aesthetic criteria that can be
recognized independently of personal attraction. In other words, we
wanted the experiment to be able to potentially capture scenarios where
a face could be judged beautiful yet not, attractive and vice versa.**

In the last part of the study, participants were informed that "about
half" of the images previously seen were AI-generated (the instructions
used a cover story explaining that the aim of the research was to
validate a new face generation algorithm). The same set of stimuli was
displayed again for 500 ms in a new randomized order. This time, after
each display, participants were asked to express their belief regarding
the nature of the stimulus using a visual analog scale (with *Fake* and
*Real* as the two extremes). The study was implemented using *jsPsych*
[@de2015jspsych], and the exact instructions are available in the
experiment code.

**Participants.** Although the main part of the study relied on
within-subject design (with 109 trials per participant), we also planned
to do between-participants analyses, thus aiming at collecting a larger
sample than traditionally used in experimental psychology (with budget
availability as the main constraint). One hundred and fifty participants
were recruited via *Prolific*, a crowd-sourcing platform recognized for
providing high quality data [@peer2022; douglas2023data]. The only
inclusion criterion was a fluent proficiency in English to ensure that
the experiment instructions would be well-understood. Participants were
incentivised with a reward of about \textsterling 7.5 for completing the
study, which took about 45 minutes to finish. Demographic variables
(age, gender, sexual orientation, education and ethnicity) were
self-reported on a voluntary basis.

We excluded 5 participants that either failed 2 (\>= 66.6%) or more
attention check questions, took an implausibly short time to finish the
questionnaires or had incomplete responses. The final sample included
145 participants (Mean age = 28.3, SD = 9.0, range: \[19, 66\]; Sex:
48.3% females, 51.0% males, 0.7% others).

**Data Analysis.** The real-fake ratings (measured originally with a
\[-1, 1\] analog scale) were converted into two scores, corresponding to
two conceptually distinct mechanisms: the dichotomous *belief* (real or
fake, based on the sign of the rating) and the *confidence* (the
rating's absolute value) associated with that belief. The former was
analyzed using logistic mixed models, which modelled the probability of
assigning a face to the real (\>= 0) as opposed to fake (\< 0). The
latter, as well as the other face ratings (attractiveness, beauty,
trustworthiness and familiarity), was modelled using mixed beta
regressions (suited for outcome variables expressed in percentages). The
models included the participants and stimuli as random factors.

We started by investigating the effect of the procedure and instructions
to check whether the stimuli (which were all images of real faces) were
judged as fake in sufficient proportion to warrant their analysis.
Additionally, we assessed the effect of the re-exposure delay, i.e., the
time between the first presentation of the image (corresponding to the
face ratings) and the second presentation (for the real-fake rating), as
well as that of the presentation order to check whether for habituation
or learning effects.

The determinants of reality beliefs were modelled separately for
attractiveness, beauty, trustworthiness, and familiarity, using second
order raw polynomials coefficients to allow for possible quadratic
relationships (**Figure 2**). Aside from attractiveness (conceptualized
as a general construct), models for beauty, trustworthiness and
familiarity were adjusted for the the two remaining variables *mutatis
mutandis*. The analysis focused on sexual-orientation relevant stimuli,
i.e., on faces that were aligned with respect to the participants'
sexual orientation (i.e., female faces for heterosexual males, male
faces for homosexual males, etc.), and the models included the
interaction with the participants' gender (as a sexual dimorphism has
been reported in face appraisal processes). For the attractiveness and
beauty models, we then added the interaction with the reported
self-attractiveness (the average of the two questions pertaining to it)
to investigate its potential modulatory effect. Finally, we investigated
the inter-individual correlates of simulation monitoring with similar
models (but this time, for all items regardless of the participant's
gender or sexual orientation) for each questionnaire, with all of the
subscales as orthogonal predictors.

```{r message=FALSE, warning=FALSE, fig.cap="Figure 2. Top part shows the efffect of face ratings on 1) the probability of judging a face as real vs. fake (solid line) and 2) on the confidence associated with that judgement (dashed lines) depending on the sex. Bottom part shows the effect of personality traits on the belief (black line) and the confidence associated with it (colored lines). The points are the average per participant confidence for both types of judgements. Stars indicate significance (p < .001***, p < .01**, p < .05*).", out.width="100%"}
knitr::include_graphics("../../figures/Figure2.png")
```

The analysis was carried out using *R 4.2* [@RCoreTeam2022], the
*tidyverse* [@wickham2019], and the *easystats* collection of packages
[@insightArticle; @bayestestRArticle; @performanceArticle;
@parametersArticle; @correlationArticle]. As all the details, scripts
and complimentary analyses are open-access, we will focus in the
manuscript on findings that are highly statistically significant
($p <.01$).

## Results

<!-- ### Manipulation Check -->

<!-- Only one image file yielded a strong simulation monitoring bias (\> 85%), being classified as fake by 88.7% of participants. This image was removed from further analysis, leaving 108 trials per participant.  -->

On average, across participants, 44% of images (95%\~CI \[0.12, 0.64\])
were judged as fake and 56% of images (95%\~CI \[0.36, 0.88\]) as real.
An intercept-only model with the participants and images as random
factors showed that the Intraclass Correlation Coefficient (ICC), which
can be interpreted as the proportion of variance explained by the random
factors, was of 9.0% for the participants and 9.6% for the stimuli.

While the delay of stimulus re-exposure stimulus did not have a
significant effect on participants' beliefs of reality ($OR = 1.00$,
$95\%~CI = [0.99, 1.00]$), judgement confidence was found to be
negatively associated with re-exposure delay when the faces were judged
as real ($\beta = -0.006$, $95\%~CI = [-0.1, 0.002]$, $p = .004$). The
presentation order also did not have have an effect on the belief
($OR = 1.00$, $95\%~CI = [1.00, 1.00]$) but was related to a decrease of
confidence ($\beta_{real} = -0.003$, $95\%~CI = [-0.004, -0.002]$,
$p < .001$; $\beta_{fake} = -0.002$, $95\%~CI = [-0.004, -0.0003]$,
$p = .021$): items presented at the end of the session were judged with
a similar bias but a decreased overall confidence.

**Determinants of Simulation Monitoring.** Attractiveness had a
significant positive and linear relationship ($R^2_{marginal}$ = 2.0%)
with the belief that a stimulus was real ($\beta_{poly1} = 16.57$,
$95\%~CI = [7.33, 25.82]$, $z = 3.51$, $p < .001$) for males, and a
quadratic relationship for females ($\beta_{poly2} = 7.82$,
$95\%~CI = [1.81, 13.84]$, $z = 2.55$, $p = .011$), with both
non-attractive and attractive faces being judged as more real.
Attractiveness was also found to have a significant positive and
quadratic relationship with confidence in judging faces both as real
($\beta_{poly2} = 4.30$, $95\%~CI = [0.97, 7.64]$, $z = 2.53$,
$p = .011$) and as fake ($\beta_{poly2} = 5.23$,
$95\%~CI = [0. 86, 9.60]$, $z = 2.35$, $p = .019$) for females. For
males, however, a significant negative and quadratic relationship was
found between attractiveness ratings and belief confidence only for
faces judged as fake ($\beta_{poly2} = -9.92$,
$95\%~CI = [-18.99, -0.86]$, $z = -2.15$, $p = .032$). There was no
interaction with reported self-attractiveness.

Beauty, adjusted for trustworthiness and familiarity, had a significant
positive and linear relationship ($R^2_{marginal}$ = 2.0%) with the
belief that a stimulus was real ($\beta_{poly1} = 11.82$,
$95\%~CI = [4.28, 20.21]$, $z = 2.76$, $p = .006$) for males only. No
effect on confidence was found, aside from a quadratic relationship in
females for faces judged as fake, suggesting that non-beautiful and
highly beautiful faces were rated as fake with more confidence than
average faces ($\beta_{poly2} = 7.84$, $95\%~CI = [3.39, 12.29]$,
$z = 3.46$, $p < .001$). There was no interaction with reported
self-attractiveness.

Trustworthiness, adjusted for beauty and familiarity, had a
predominantly positive and linear relationship ($R^2_{marginal}$ = 2.0%)
with the belief that a stimulus was real ($\beta_{poly1} = 6.44$,
$95\%~CI = [-0.11, 13.00]$, $z = 1.93$, $p = .0054$) for females only.
No effect on confidence was found for males, whereas a quadratic
relationship was found for females for both faces judged as real
($\beta_{poly2} = 6.14$, $95\%~CI = [2.13, 10.14]$, $z = 3.00$,
$p = .003$) as well as fake ($\beta_{poly2} =6.12$,
$95\%~CI = [1.49, 10.75]$, $z = 2.59$, $p = .001$), suggesting that
non-trustworthy and highly trustworthy faces were rated with more
confidence than average faces.

We did not find any significant relationships for familiarity adjusted
for beauty and trustworthiness ($R^2_{marginal}$ = 2.0%). However, a
significant positive and linear relationship was found between
familiarity and the confidence judgements of rating faces as real
($\beta_{poly1} = 9.98$, $95\%~CI = [3.83, 16.13]$, $z = 3.18$,
$p = .001$) whereas a negative linear relationship was found with those
judged as fake ($\beta_{poly1} = -12.41$, $95\%~CI = [-20.27, -4.54]$,
$z = -3.09$, $p = .002$) for males only. This hence suggests that males
more confidently judge faces as real with when they are familiar, and as
fake when they are unfamiliar.

Note that we also tested as predictors the normative attractiveness and
trustworthiness scores (i.e., the average values from the stimuli
database validation), which showed a significant positive linear
relationship between beliefs of reality and attractiveness, as well as
trustworthiness, only for males (see Supplementary Analysis for
details).

**Inter-Individual Correlates of Simulation Monitoring.** The models
including the personality traits suggested that *Honesty-Humility* had a
significant negative relationship with the confidence associated with
real as well as fake judgements ($\beta_{real} = -1.62$,
$95\%~CI = [-2.55, -0.70]$, $z = -3.43$, $p < .001$;
$\beta_{fake} = -1.16$, $95\%~CI = [-2.09, -0.23]$, $z = -2.45$,
$p = 0.014$).

Significant positive associations were found between the probability of
judging faces as real and dimensions of narcissism such as *Acclaim
Seeking* ($\beta = 2.24$, $95\%~CI = [1.17, 4.27]$, $z = 2.44$,
$p = .015$), and *Manipulativeness* ($\beta = 0.47$,
$95\%~CI = [0.25, 0.87]$, $z = -2.4$, $p = 0.017$). Confidence
judgements also shared significant links with narcissism through various
facets, such as a positive relationship between the confidence for both
real and fake judgements with *Acclaim Seeking* ($\beta_{real} = 1.65$,
$95\%~CI = [0.59, 2.70]$, $z = 3.07$, $p =.002$; $\beta_{fake} = 1.62$,
$95\%~CI = [0.56, 2.68]$, $z = 3.00$, $p = .003$), and a negative
relationship with *Authoritativeness* ($\beta_{real} = -1.57$,
$95\%~CI = [-2.58, -0.57]$, $z = -3.08$, $p = .002$;
$\beta_{fake} = -1.49$, $95\%~CI = [-2.50, -0.48]$, $z = -2.89$,
$p = .004$).

A positive trend was found in the relationship between the *Persecutory
Ideation* dimension of paranoid thinking and the belief that the faces
were real ($\beta = 1.87$, $95\%~CI = [0.99, 3.54]$, $z = 1.93$,
$p = .054$).

The *Prospective Anxiety* aspect of intolerance to uncertainty shared a
negative trend in its association with confidence ratings
($\beta_{real} = 1.43$, $95\%~CI = [0.10, 2.76]$, $z = 2.10$,
$p = .036$; $\beta_{fake} = -0.91$, $95\%~CI = [-1.93, 0.11]$,
$z = -1.75$, $p = .081$). No significant effect was found for social
anxiety.

Questions pertaining to the attitude towards AI were reduced to 3
dimensions through factor analysis, labelled AI-Enthusiasm (loaded by
items expressing interest and excitement in AI development and
applications), AI-Realness (loaded by items expressing positive opinions
on the ability of AI to create realistic material), and AI-Danger
(loaded by items expressing concerns on the unethical misuse of AI
technology). However, only AI-Enthusiasm displayed a significant
positive relationship with the confidence in both real and fake
judgements ($\beta_{real} = 0.21$, $95\%~CI = [0.02, 0.40]$, $z = 2.20$,
$p = .028$; $\beta_{fake} = 0.31$, $95\%~CI = [0.12, 0.50]$,
$z = -8.90$, $p < 0.001$).

## Discussion

This study aimed at investigating the effect of facial ratings
(attractiveness, beauty, trustworthiness and familiarity) on simulation
monitoring, i.e., on the belief that a stimulus was artificially
generated. Most strikingly, despite all the stimuli being real faces
from the same database, all participants believed (to high degrees of
confidence) that a significant proportion of them were fake. This
finding not only attests to the effectiveness of our instructions, but
highlights the current levels of expectation regarding CGI technology.
The strong impact of prior expectations and information on reality
beliefs underlines the volatility of our sense of reality. In fact,
stimuli-related and participant-related characteristics accounted
together for less than 20% of the beliefs variance, suggesting a large
contribution of other subjective processes.

<!-- Attractiveness: females -->

Although attractiveness did not seem to be the primary drive underlying
simulation monitoring of face images, we do nonetheless report
significant associations, with different patterns observed depending on
the participant's gender. The quadratic relationship found for female
participants is aligned with our hypothesis that salient faces (i.e.,
rated as very attractive or very unattractive) are judged to be more
real. The fact that this effect did not reach significance for beauty
underlines that attractiveness judgement, and its role in simulation
monitoring, is a multidimensional construct that cannot be reduced to
physical facial attractiveness, in particular for women [@buunk2002age;
@qi2022gender]. In fact, female participants were more confident in
judging faces as fake only when they were rated very high or low on
beauty, suggesting that physical beauty and attractiveness are not
analogous in their effects on simulation monitoring decisions.

<!-- Attractiveness: males -->

Interestingly, we found a significant positive linear relationship in
male participants for both attractiveness and beauty on simulation
monitoring that we could interpret under an evolutionary lens.
Specifically, males purportedly place more emphasis on facial
attractiveness as a sign of reproductive potential, as compared with
females, who tend to value characteristics signaling resource
acquisition capabilities [@qi2022gender; @fink2006facial;
@buunk2002age]. It is thus possible that the evolutionary weight
associated with attractiveness skewed the perceived saliency of
attractive faces for men, rendering them significantly more salient than
unattractive faces, and in turn distorting the relationship with
simulation monitoring.
<!-- However, it is also possible that this pattern occurred due to the stimuli range, by which the stimuli did not include sufficiently unattractive faces for males to be , by which   -->
However, future studies should test this saliency-based hypothesis by
measuring constructs closer to salience and its effects, for instance
using neuroimaging [@lou2015perceptual; @indovina2007dissociation] or
physiological markers
<!-- (e.g., heart rate deceleration) [@skora2022functional]. --> [e.g.,
heart rate deceleration, @skora2022functional].

<!-- trustworthiness -->

Our findings do not support the existence of a strong link between
perceived trustworthiness and reality judgments. Given prior evidence
that faces presented as computer-generated were rated less trustworthy
[@liefooghe2022faces; @hoogers2021effect; @balas2017], we expected such
a linear association to be more clearly present. However, our results
suggest a relationship with confidence ratings, especially for women,
whereby faces judged with low and high trustworthiness are judged as
real and fake with higher confidence. One of the underlying mechanisms
that possibly contributed to this dimorphism could be the increased
risk-taking aversion reported in females [explained evolutionarily as a
compromise to their reproductive potential, @van2020sex], to which
perceived facial trustworthiness relates [@hou2019survival]. Future
studies should clarify the role of trustworthiness both as a predictor
and outcome of reality beliefs.

<!-- However, if that was the case, faces judged as highly untrustworthy should have appeared as even more salient (representing an evolutionary threat), and hence be judged as more real, leading to a quadratic relationship between trustworthiness and simulation monitoring instead. Further studies are needed to investigate the causes of the increased simulation monitoring sensitivity to trustworthiness in females. -->

<!-- familiarity -->

Contrary to our hypothesis, we did not find familiarity to be
significantly related to simulation monitoring decisions. Interestingly,
there were significant linear relationships between familiarity and
confidence judgements for males only, where familiarity increased the
confidence of reality beliefs. Although the familiarity measure was not
a "recognition" measure, evidence from studies pertaining to the latter
could be linked, reporting better face memory for females
[@mishra2019gender; @lewin2002sex; @sommer2013sex], as well as an
overconfidence in face recall for males [@bailey2021gender;
@herbst2020gender]. However, it should be noted that the distribution of
familiarity ratings was strongly skewed, and only a low number of
pictures was rated as highly familiar. As such, future studies should
clarify this point by experimentally manipulating familiarity, for
instance by modulating the amount of exposure to items before querying
the simulation monitoring judgements.

<!-- inter-indiv correlates -->

Regarding the role of inter-individual characteristics in simulation
monitoring tendencies, we found higher scores of honesty-humility - a
trait related to an increased risk perception and aversion
[@levidi2022understanding; @weller2012honest] - to be related to a lower
confidence in simulation monitoring judgements. Notably, greater
narcissistic tendencies in dimensions such as acclaim seeking were
associated with a higher number of faces judged as real. This is in line
with recent research which found people with higher narcissism scores
less likely to engage in analytical reasoning strategies such as
reflective thinking [@littrell2020overconfidently;
@ahadzadeh2021social], and to be more vigilant and attentive to external
stimuli [@carolan2017searching; @grapsas2020and; @eddy2021self].

Moreover, putting the significant positive links between narcissistic
acclaim seeking and confidence judgements in perspective with the
negative correlation between honesty-humility and narcissism
[@hodson2018dark], we confirm previous evidence regarding the
relationship between narcissistic grandiosity and over-confidence in
decision-making [@campbell2004narcissism; @chatterjee2017master;
@brunell2017narcissism; @o2021grandiose]. Although an inverse effect was
found for the narcissistic facet of authoritativeness, we interpret this
relationship as related to a higher response assertiveness. Taken
together, these results suggest that participants with low humility and
high recognition desires are more confident in their judgement regarding
the real or fake nature of ambiguous stimuli. Alternatively,
participants with opposite traits might perceive a higher risk in the
decision-making process and its potential consequences (e.g., being seen
as bad at the task at hand), resulting in more conservative confidence
ratings.

<!-- Social Anxiety, Paranoia and Uncertainty Intolerance -->

Our findings suggest - though with low certainty - a potential positive
link between paranoid ideation and the tendency to believe that the
stimuli were real. Given previous reports that people with higher levels
of paranoia are more sensitive to cues of social threat
[@fornells2015people; @freeman2003can; @king2017paranoia], it is
plausible that paranoid traits confer greater saliency and emotionality
to observed faces, hence increasing perceptions of its realness. This
hypothesis, if confirmed by future studies, would be in line with
previous findings that persecutory delusions are predicted by a greater
sense of presence in VR environments populated with virtual characters
[@freeman2005psychology].

<!-- AI attitudes -->

Despite the ubiquity of AI, the literature pertaining to the influence
of people's AI attitudes on simulation monitoring is scarce. Contrary to
our expectations, we did not find evidence for the role of participants'
expectations regarding the capabilities of AI technology (in terms of
the realism of its productions). Instead, we found only one's enthusiasm
about AI technology to be related to an increased confidence in
simulation monitoring ratings. This could potentially be because
participants with a highly positive attitude towards AI perceive
themselves as having greater knowledge about AI and its capabilities
[@said2022artificial], hence permitting themselves to be more confident
in their simulation monitoring decisions. In fact, this result is in
line with reports that AI attitudes interacts with people's perceived
self-knowledge to influence their perception of the opportunities and
risks accorded by AI applications [@said2022artificial].

<!-- Re-exposure Delay -->

On a methodological level, although the order of presentation of the
facial images was randomized to reduce effects of adaptation,
participants were more confident in their judgements for faces perceived
as real following a shorter re-exposure delay. Such shorter durations
could be associated with the faces being better remembered and appearing
more familiar, thereby triggering self-referential and autobiographical
memory processing during the repeated display [@abraham2009reality;
@gobbini2013prioritized; @taylor2009neural]. Indeed, this finding is
consistent with studies in which fictional stimuli that were associated
with familiarity up-regulated emotions, biasing its salience and
perceived realness [@sperduti2016paradox; @makowski2017being]. However,
if that was the case, we would expect shorter re-exposure delays to
impact the decision bias as well towards reality, rather than simply the
confidence. Future studies should further investigate the modulatory
effects of types and degrees of familiarity on perceived realness
judgements.

<!-- Limitations -->

Several limitations have to be noted. The current experimental paradigm
required participants to judge the realness of faces they had prior
exposure to (which was done to prevent reality judgements from
influencing the other ratings). Although the effect of re-exposure delay
was negligible, the potential bias induced by face familiarity (as
compared to judging completely new items) cannot be discarded. Future
studies could examine that by incorporating novel face images or
increasing the duration of the re-exposure delay.

Another issue is the impact on reality judgements of the prior explicit
instruction that "about half of the faces were AI-generated and the
other half real photos". Given this prior information given to
participants, it might seems like our enthusiasm pertaining to the
finding that most people did indeed believe a high number of stimuli to
be fake might be unwarranted, since it simply affirms participants
followed the instructions. However, even if that was the case, the
finding that our beliefs of reality can be so easily re-programmed with
simple instructions and lead to high-confidence answers remains an
interesting phenomenon. Moreover, it is to note that the paradigm did
not instruct participants to balance their answers according to a
certain distribution (e.g., 50-50), merely providing them a description
of the dataset. The fact that no presentation order effect was found on
reality beliefs suggests that participants did not try to actively
distribute their responses to match the instructions, in which case we
would have expected a different pattern: for instance, the first few
items judged as real (the initial "true" belief of the participants),
and a bias would progressively appear towards responding "fake" (as
participants realize that all stimuli are of similar nature and that
they have to "make up" for the prevalence of their "real" answers to
fulfil the expected proportion of responses given the instructions).

That said, the potential demand effect of the instructions still exists,
and a control condition without the cover story with AI-generated images
would in-principle be able to mitigate such confounds to some extent.
However, the distinction real/fake is hard to operationalize and
introduce to participants in a vacuum (simply instructing them to
discriminate real from fake without providing some background
information regarding the context and defining what is meant by "fake"
seems hardly feasible). That being the case, future studies should study
the impact of these higher-order expectations on ratings [for instance,
@tucciarelli2020 found that merely mentioning that some faces were AI
generated decreased, on average, the trustworthiness ratings for all
faces] as well as on the simulation monitoring process itself (i.e., the
"criterion": would people form and distribute judgements differently).
This can be studied by modulating this expectation in a controlled
fashion (e.g., "most of the images *but a few* are real" vs. "most of
the images *but a few* are fake") or inventing some implicit way of
measuring reality belief that would not require the explicit
introduction of the concept of fake vs. real to participants.

Finally, it is important to note that although consistent in their
directions across models and variables, the magnitude of the effects
found in the study was relatively small, suggesting that the facial
appraisals measured in the study were not the key determinants of
simulation monitoring. Hence, beyond exploring new potential mechanisms,
future studies should include a more thorough debriefing to try to
capture what conscious strategies (if any) the participants used (e.g.,
focusing on some features of the stimulus - like hair or eyes in the
case of faces) to guide their reality beliefs.

<!-- Conclusion -->

In summary, the aim of the present study was to examine whether a subset
of specific characteristics, in particular face attractiveness,
significantly influences our simulation monitoring decisions. Notably,
we found faces rated as attractive to be perceived as more real, with a
possible sexual dimorphism affecting the shape of the relationship. We
also found that inter-individual traits, such as narcissistic
acclaim-seeking, honesty-humility, and paranoid ideation, were related
to a systematic bias towards beliefs that the stimuli were real or fake.
We believe that these findings provide the foundations to help us
understand what drives reality beliefs in an increasingly
reality-ambiguous world.

# Data Availability

The datasets generated and/or analysed during the current study are
available in the GitHub repository
https://github.com/RealityBending/FakeFace

\pagebreak

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::